figures and calculations for global fisheries management

### 2019-01-15 documents/figs-and-calcs.Rmd
added in cooperatives, will check results in the morning

projection_data_line65 <- readRDS(here("data","projection_data.rds")) %>% 
  filter(Scenario == "Historic") %>%
  group_by(IdOrig) %>% 
  filter(Year == max(Year))%>%
  filter(Dbase == "FAO") %>%
  filter(IdLevel != "Neis")%>%
  filter(CatchShare == "1")

@@

turf_catch_lower <- projection_corb%>%
  filter(turf == "TRUE")
sum(turf_catch_lower$Catch)
#301,000 lower limit

turf_catch_upper <- projection_corb_2%>%
  filter(turf == "TRUE")
sum(turf_catch_upper$Catch)
#1,370,210 upper limit
#upper bound of total percent of global catch from turf:
1370210/80000000
#0.01712763

###cooperative estimates
coop_catch <- projection_corb%>%
  filter(coop == "TRUE")
sum(coop_catch$Catch)
#443231
#percent of global catch from coop:
443231/80000000
#0.005540387

@@

@@

@@

#GDP data
gdp_all <- read_excel(here("data","un_gdp_2016.xls"))

#EDF TURF data
edf_turf <- read_excel(here("data/EDF_TURF_data.xlsx"))%>%
  select("Country", "SciName","CatchShare") %>%
  filter(SciName != "NA")

#DiscoverTURFs data
d_turfs <- read_excel(here("data/DiscoverTURFs_simple.xlsx")) %>%
  select("Country", "SciName", "CatchShare") %>%
  filter(SciName != "NA")

@@

all_turfs <- dplyr::full_join(d_turfs, edf_turf, by= c("Country", "SciName", "CatchShare")) %>%
  distinct()

@@

#add the turf data to upside
projection_turfs <- full_join(projection_data, turfs, by = c("Country", "SciName", "SpeciesCat","Year"), all = T) %>% 
  mutate(turf = ifelse(is.na(turf),0,1)) %>% 
  mutate(CatchShare = ifelse(turf == 1,2, CatchShare))

@@

#question here, will the filter remove out when year is NA? the turf data doesn't have this resolution; edf does have start date though
catch_shares <- left_join(projection_turfs, gdp, by = "Country") %>% 
  group_by(IdOrig) %>%
  dplyr::mutate(max_year = max(Year)) %>% 
  filter(max_year== 2012,
         CatchShare %in% c(0,1),
         !is.na(gdp_center),
         !is.na(MainCat),
         Year == max_year) %>% 
  ungroup() %>% 
  mutate(MainCat = factor(MainCat))

catch_shares_and_turfs <- left_join(projection_turfs, gdp, by = "Country") %>% 
  group_by(IdOrig) %>%
  dplyr::mutate(max_year = max(Year)) %>% 
  filter(max_year== 2012,
         CatchShare %in% c(1,2),
          !is.na(gdp_center),
         Year == max_year) %>% 
  ungroup() %>% 
  mutate(MainCat = factor(MainCat)) %>% 
  mutate(CatchShare = CatchShare - 1)

everything <- left_join(projection_turfs, gdp, by = "Country") %>% 
  group_by(IdOrig) %>%
  dplyr::mutate(max_year = max(Year)) %>% 
  filter(max_year== 2012,
         Year == max_year,
                   !is.na(gdp_center)) %>% 
  ungroup() %>% 
  mutate(MainCat = factor(MainCat))

@@

upjoin <-
  caret::upSample(
    x = select(catch_shares_and_turfs, -CatchShare),
    y = factor(catch_shares_and_turfs$CatchShare),
    yname = "CatchShare"
  )

cs_or_turf_fit <-
  glm(
    formula = CatchShare ~ gdp_center + SpeciesCatName,
    family = "binomial",
    data = upjoin
  )

summary(cs_or_turf_fit)

@@

  cs_or_turf_pred_plot <- orig_cs_or_turf_fit %>%
  ggplot(aes(pred, CatchShare == 1, fill = CatchShare == 1)) +
  ggridges::geom_density_ridges(show.legend = FALSE, alpha = 0.75) +
  geom_vline(aes(xintercept = 0.5), linetype = 2, color = "red") +
  labs(y = "Is TURF?",
       x = "Probability of TURF",
       caption = "Classified as TURF if p>0.5") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, .01))

g <- pROC::roc(CatchShare ~ pred, data = orig_cs_or_turf_fit)

@@

- “current annual net benefits from marine capture fisheries are tens of billions of U.S. dollars less than the potential benefits”
```{r}
include_graphics("sunkenbillionstable.png")
```

@@

>>>>

projection_data_line65 <- readRDS(here("data","projection_data.rds")) %>% 
  filter(Scenario == "Historic") %>%
  group_by(IdOrig) %>% 
  filter(Year == 2012) %>%
  filter(CatchShare == 1)

@@

turf_catch_lower <- projection_corb%>%
  filter(turf == TRUE)
sum(turf_catch_lower$Catch)
#301,000 lower limit

turf_catch_upper <- projection_corb_2%>%
  filter(turf == TRUE)
scales::percent(sum(turf_catch_upper$Catch) / 80e6, 1)
#1,370,210 upper limit
#upper bound of total percent of global catch from turf:
1370210/80000000
#0.01712763

###cooperative estimates
coop_catch <- projection_corb%>%
  filter(coop == "TRUE")
(sum(coop_catch$Catch)/ 80e6) %>% scales::percent(1)
#443231
#percent of global catch from coop:
443231/80000000
#0.005540387

@@

## line 69

based on this calculation, we find that

a= projection_data %>% 
  filter(Year == 2012) %>% 
  group_by(Country, CatchShare) %>% 
  count()

scales::percent(mean(a$CatchShare),1)
```

@@

```{r prepare-data}

@@

taxa_lookup <- projection_data %>% 
  select(SciName, SpeciesCat, SpeciesCatName) %>% unique()

projection_data$clean_country <- countrycode::countrycode(projection_data$Country,"country.name","country.name")

#GDP data
gdp <- read_excel(here("data","un_gdp_2016.xls")) %>% 
  mutate(clean_country = countrycode::countrycode(Country,"country.name","country.name"))

#EDF TURF data
edf_turf <- read_excel(here("data/EDF_TURF_data.xlsx")) %>%
  select("Country", "SciName", "CatchShare") %>%
  filter(SciName != "NA") %>%
  mutate(clean_country = countrycode::countrycode(Country, "country.name", "country.name"))

#DiscoverTURFs data
d_turfs <- read_excel(here("data/DiscoverTURFs_simple.xlsx")) %>%
  select("Country", "SciName", "CatchShare") %>%
  filter(SciName != "NA") %>% 
    mutate(clean_country = countrycode::countrycode(Country,"country.name","country.name"))

@@

all_turfs <- dplyr::full_join(d_turfs, edf_turf, by= c("clean_country", "SciName", "CatchShare")) %>%
  distinct()

@@

cooperatives <- read_csv(here("data","clean_cooperatives_data.csv"))

coop_scinames <- taxize::comm2sci(unique(cooperatives$target_species))

coops_names <- tibble(target_species = unique(cooperatives$target_species), SciName =  coop_scinames)  %>% 
  unnest()

coops <- cooperatives %>% 
  select(fishery_id, host_country, target_species) %>% 
  left_join(coops_names, by = "target_species") %>% 
  left_join(taxa_lookup, by = c("SciName" = "SciName")) %>% filter(!is.na(SpeciesCat)) %>% 
     mutate(clean_country = countrycode::countrycode(host_country,"country.name","country.name")) %>% 
  mutate(cooperative = 1)

projection_data %>% 
  filter(Year == 2012) %>% 
  group_by(CatchShare) %>% 
  summarise(tc = sum(Catch, na.rm = T))

#add the turf data to upside
projection_turfs <- full_join(projection_data, turfs, by = c("clean_country", "SciName", "SpeciesCat","Year"), all = T) %>% 
  mutate(turf = ifelse(is.na(turf),0,1)) %>% 
  mutate(CatchShare = ifelse(turf == 1,2, CatchShare)) %>%
  left_join(coops %>% select(SciName, clean_country, cooperative), by = c("SciName", "clean_country")) %>% 
  left_join(gdp, by = "clean_country") %>% 
    mutate(turf = ifelse(cooperative == 1,1,turf))

projection_turfs %>% 
  filter(Year == 2012) %>% 
  group_by(CatchShare) %>% 
  summarise(tc = sum(Catch, na.rm = T))

@@

# gdp <- gdp_all %>%
#   select(Country, gdp_center) %>%
#   filter( gdp_center != "NA")

#question here, will the filter remove out when year is NA? the turf data doesn't have this resolution; edf does have start date though
catch_shares <- projection_turfs %>% 
  group_by(IdOrig) %>%
  dplyr::mutate(max_year = max(Year)) %>% 
  filter(max_year== 2012,
         CatchShare %in% c(0,1),
         !is.na(gdp_center),
         !is.na(MainCat),
         Year == max_year) %>% 
  ungroup() %>% 
  mutate(MainCat = factor(MainCat))

catch_shares_and_turfs <- projection_turfs %>% 
  group_by(IdOrig) %>%
  dplyr::mutate(max_year = max(Year)) %>% 
  filter(max_year== 2012,
         CatchShare %in% c(1,2),
          !is.na(gdp_center),
         Year == max_year) %>% 
  ungroup() %>% 
  mutate(MainCat = factor(MainCat)) %>% 
  mutate(CatchShare = CatchShare - 1) %>% 
  select(-gdp) %>% 
  rename(TURF = CatchShare,
         gdp = gdp_center)

everything <- projection_turfs %>% 
  group_by(IdOrig) %>%
  dplyr::mutate(max_year = max(Year)) %>% 
  filter(max_year== 2012,
         Year == max_year,
                   !is.na(gdp_center)) %>% 
  ungroup() %>% 
  mutate(MainCat = factor(MainCat))

@@

catch_shares_and_turfs %>% 
  filter(Year == 2012) %>% 
group_by(TURF) %>% 
  summarise(n = length(Catch),
            tc = sum(Catch, na.rm = T))
upjoin <-
  caret::upSample(
    x = select(catch_shares_and_turfs, -TURF),
    y = factor(catch_shares_and_turfs$TURF),
    yname = "TURF"
  )

cs_or_turf_fit <-
  glm(
    formula = TURF ~ gdp + SpeciesCatName,
    family = "binomial",
    data = catch_shares_and_turfs 
  )

covar_names <-  broom::tidy(cs_or_turf_fit)$term %>% 
  str_replace_all("SpeciesCatName","")

stargazer::stargazer(cs_or_turf_fit,
                     covariate.labels =covar_names,
                     intercept.top = TRUE,
                     intercept.bottom = FALSE)

@@

temp <- projection_turfs %>%
  filter(SpeciesCatName %in% unique(catch_shares_and_turfs$SpeciesCatName),
         Year == 2012) %>% 
  select(-gdp) %>% 
  rename(gdp = gdp_center)

global_cs_or_turf_fit <- temp %>%
  ungroup() %>%
  mutate(pred = 1 - predict(cs_or_turf_fit, type = "response", newdata = temp)) %>% 
  mutate(rbfm_type = ifelse(pred > 0.5, "ITQ","TURF"))

rbfm_pred_summary <- global_cs_or_turf_fit %>% 
  filter(!is.na(rbfm_type), CatchShare == 0) %>% 
  group_by(rbfm_type) %>% 
  summarise(n = length(MSY),
            tmsy = sum(MSY, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(pn = n / sum(n),
         pmsy = tmsy / sum(tmsy))

country_pred_summary <- global_cs_or_turf_fit %>%
  filter(!is.na(rbfm_type), CatchShare == 0,!is.na(clean_country)) %>%
  group_by(clean_country, rbfm_type) %>%
  summarise(n = length(MSY),
            tmsy = sum(MSY, na.rm = TRUE)) %>%
  group_by(clean_country) %>%
  mutate(pn = n / sum(n),
         pmsy = tmsy / sum(tmsy)) %>%
  mutate(ttmsy = sum(tmsy)) %>% 
  ungroup() %>% 
  select(clean_country,ttmsy, rbfm_type,pmsy) %>% 
  spread(rbfm_type,pmsy, fill = 0) %>% 
  arrange(desc(ttmsy)) %>% 
  slice(1:20) %>% 
  mutate(`MSY Rank` = 1:nrow(.)) %>% 
  select(-ttmsy) %>% 
  rename("Country" = "clean_country") %>% 
  select(Country, `MSY Rank`, ITQ,TURF) %>% 
  mutate(ITQ = scales::percent(ITQ,1),
         TURF = scales::percent(TURF,1))

knitr::kable(country_pred_summary, format = "latex", longtable = TRUE)

  cs_or_turf_pred_plot <- orig_cs_or_turf_fit %>%
  ggplot(aes(pred, TURF == 1, fill = TURF == 1)) +
  ggridges::geom_density_ridges(show.legend = FALSE, alpha = 0.75) +
  geom_vline(aes(xintercept = 0.5), linetype = 2, color = "red") +
  labs(y = "Is TURF?",
       x = "Probability of TURF",
       caption = "Classified as TURF if p>0.5") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, .01))

g <- pROC::roc(TURF ~ pred, data = orig_cs_or_turf_fit)

@@

- “current annual net benefits from marine capture fisheries are tens of billions of U.S. dollars less than the potential benefits”
```{r}
knitr::include_graphics("sunkenbillionstable.png")
```

@@

## line 249

recent <- projection_data %>% 
  filter(Year == 2012,
         Scenario == "Historic")

dim(recent)

sum(recent$Catch)

62399651 / 80e6
